---
title: "Homework Q6"
author: "Gayatri Mainkar"
date: "1/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The bio3d package needs to be loaded in order to use this function.

```{r}
library("bio3d")
```

The below code chunk contains a function that analyzes protein drug interactions by reading in any protein PDB data and outputting a plot of the B-factor trend for the specified protein. Simply input the PDB identifier of the protein to obtain this plot.

Input: a single element character vector containing the four letter PDB identifier for online file access.

Output: Plot with the B-factor trend of the protein.

```{r}
btrend <- function(x){
  s <- read.pdb(x) #reads in specified protein
  s.chainA <- trim.pdb(s, chain="A", elety="CA") #trims protein to subset of its atoms in chain A of type "CA"
  s.b <- s.chainA$atom$b #assigns vector of b-factors of each atom to variable
  plotb3(s.b, sse=s.chainA, typ="l", ylab="Bfactor", main = x) #plots the trend of the b-factor
}
```

Test Case #1
```{r}
btrend("4AKE")
```

Test Case #2
```{r}
btrend("1AKE")
```

Test Case #3
```{r}
btrend("1E4Y")
```

